---
title: "Urchin_Practice"
format: html
editor: visual
---

## Lab Feeding Trials

no_choice \<-

```{r}
#| label: installing-packages

library(tidyverse)
library(dplyr)
library(ggplot2)
library(janitor)

no_choice_raw <- read_csv("data/no_choice_raw.csv")

no_choice <- clean_names(no_choice_raw)
```

```{r}
no_choice |>
  ggplot(aes(x = treatment, y = amount_eaten, fill = treatment)) +
  geom_boxplot(alpha = 0.75) +
  labs(
    y = "Amount Consumed (g)",
    x = NULL
  ) +
  scale_fill_manual(values = c(
    detrital = "#354230",
    halodule = "#006600",
    zostera = "#9ACD23"
  )) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}

anova_model <- aov(amount_eaten ~ treatment, data = no_choice)
summary(anova_model)

TukeyHSD(anova_model)

list(
  ANOVA = data.frame(
    Source    = c("treatment", "Residuals"),
    Df        = c(2, 33),
    Sum_Sq    = c(5.56, 20.71),
    Mean_Sq   = c(2.78, 0.63),
    F_value   = c(4.429, NA),
    P_value   = c(0.0198, NA)
  ),
  Tukey = data.frame(
    Comparison   = c("halodule vs detrital", "zostera vs detrital", "zostera vs halodule"),
    Diff         = c(-0.7717, -0.8842, -0.1125),
    P_value      = c(0.0580, 0.0263, 0.9356),
    Significance = c("ns", "*", "ns")
  )
)

plot(anova_model)
```

```{r}
#kruskal willis analysis -- rank based test alternative to anova bc less sensitive to assumtion
```

```{r}

choice_raw <- read_csv("data/choice_trials_raw.csv")

choice_data <- clean_names(choice_raw)
```

```{r}

choice_data <- choice_data |>
  mutate(pair = paste(choice_a, "vs", choice_b))

```

```{r}
choice_data |>
  pivot_longer(cols = c(a_amount_eaten, b_amount_eaten),
               names_to = "species_slot",
               values_to = "amount_eaten") |>
  mutate(species = ifelse(species_slot == "a_amount_eaten", choice_a, choice_b)) |>
  group_by(pair, species) |>
  summarise(mean = mean(amount_eaten, na.rm = T),
            se   = sd(amount_eaten, na.rm = T) / sqrt(n()),
            .groups = "drop") |>
  ggplot(aes(x = species, y = mean, fill = species)) +
    geom_col(color = "black", width = 0.6) +
    geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
    facet_wrap(~ pair, scales = "free_x") +
    labs(y = "Amount eaten (mg)", x = NULL) +
    scale_fill_manual(values = c(
      detrital = "#354230",
      halodule = "#006600",
      zostera = "#9ACD23"
    )) +
    theme_classic() +
    theme(panel.spacing = unit(2, "lines"))
```

```{r}
choice_data |>
  group_by(pair) |>
  summarise(p_value = t.test(a_amount_eaten, b_amount_eaten, paired = TRUE)$p.value,
            .groups = "drop")
```

report \^ this one because for proportions the others are already consumed

```{r}
choice_data |>
  mutate(total = a_amount_eaten + b_amount_eaten,
         a_pct = 100 * a_amount_eaten / total,
         b_pct = 100 * b_amount_eaten / total) |>
  group_by(pair) |>
  summarise(p_value = t.test(a_pct, b_pct, paired = TRUE)$p.value,
            .groups = "drop")
```

```{r}
choice_data |>
  filter(a_amount_eaten >= 0, b_amount_eaten >= 0) |>
  mutate(total = a_amount_eaten + b_amount_eaten,
         a_pct = 100 * a_amount_eaten / total,
         b_pct = 100 * b_amount_eaten / total) |>
  pivot_longer(cols = c(a_pct, b_pct),
               names_to = "species_slot",
               values_to = "pct_eaten") |>
  mutate(species = ifelse(species_slot == "a_pct", choice_a, choice_b)) |>
  group_by(pair, species) |>
  summarise(mean = mean(pct_eaten, na.rm = T),
            se   = sd(pct_eaten, na.rm = T) / sqrt(n()),
            .groups = "drop") |>
  ggplot(aes(x = species, y = mean, fill = species)) +
    geom_col(color = "black", width = 0.6) +
    geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
    facet_wrap(~ pair, scales = "free_x") +
    labs(y = "Percent eaten (%)", x = NULL) +
    scale_fill_manual(values = c(
      detrital = "#354230",
      halodule = "#006600",
      zostera = "#9ACD23"
    )) +
    theme_classic() +
     theme(panel.spacing = unit(2, "lines"))
```

```{r}
choice_data |>
  mutate(total = a_amount_eaten + b_amount_eaten,
         a_pct = 100 * a_amount_eaten / total,
         b_pct = 100 * b_amount_eaten / total) |>
  group_by(pair) |>
  summarise(p_value = t.test(a_pct, b_pct, paired = TRUE)$p.value,
            .groups = "drop") |>
  mutate(significance = case_when(
    p_value < 0.001 ~ "***",
    p_value < 0.01  ~ "**",
    p_value < 0.05  ~ "*",
    TRUE            ~ "ns"
  ))
```

```{r}
# non-adjusted values

choice_data |>
  mutate(
    a_amount_eaten = a_before - a_after,
    b_amount_eaten = b_before - b_after
  ) |>
  mutate(pair = paste(choice_a, "vs", choice_b)) |>
  pivot_longer(cols = c(a_amount_eaten, b_amount_eaten),
               names_to = "species_slot",
               values_to = "amount_eaten") |>
  mutate(species = ifelse(species_slot == "a_amount_eaten", choice_a, choice_b)) |>
  group_by(pair, species) |>
  summarise(mean = mean(amount_eaten, na.rm = T),
            se   = sd(amount_eaten, na.rm = T) / sqrt(n()),
            .groups = "drop") |>
ggplot(aes(x = species, y = mean, fill = species)) +
  geom_col(color = "black", width = 0.6) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2) +
  facet_wrap(~ pair, scales = "free_x") +
  labs(y = "Amount eaten (mg)", x = NULL) +
  scale_fill_manual(values = c(
    detrital = "#354230",
    halodule = "#006600",
    zostera  = "#9ACD23"
  )) +
  theme_classic() +
  theme(panel.spacing = unit(2, "lines"))
```

```{r}

# detrital vs halodule
dvh <- choice_data |> filter(choice_a == "detrital", choice_b == "halodule") |>
  mutate(a_eaten = a_before - a_after, b_eaten = b_before - b_after)
t.test(dvh$a_eaten, dvh$b_eaten, paired = TRUE)

# detrital vs zostera
dvz <- choice_data |> filter(choice_a == "detrital", choice_b == "zostera") |>
  mutate(a_eaten = a_before - a_after, b_eaten = b_before - b_after)
t.test(dvz$a_eaten, dvz$b_eaten, paired = TRUE)

# zostera vs halodule
zvh <- choice_data |> filter(choice_a == "zostera", choice_b == "halodule") |>
  mutate(a_eaten = a_before - a_after, b_eaten = b_before - b_after)
t.test(zvh$a_eaten, zvh$b_eaten, paired = TRUE)
```
